\documentclass[a4paper]{article}

\usepackage[T1]{fontenc}
\usepackage{xcolor}
\definecolor{DarkRed}{rgb}{0.6,0,0.1}

\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{amsmath, amsthm, amssymb}
\usepackage[colorlinks=true, linkcolor=DarkRed]{hyperref}
\usepackage{listings}

\setlength\parindent{0pt}
\setlength\parskip{2pt}

\definecolor{DarkBlue}{rgb}{0,0,0.5}

\lstset {
    language=C++,
    %backgroundcolor=\color{black!5}, % set backgroundcolor
    basicstyle=\small\ttfamily,
    keywordstyle=\bf\color{DarkBlue},
    commentstyle=\color{magenta},
    otherkeywords={Type,Type1,Type2,Type3,constexpr},
    keepspaces=true,
    showspaces=true,
}

\newcommand{\inlinecode}[1]{\lstinline|#1|}
%\newcommand{\TODO}[1]{{\color{red}\textbf{#1}}}

\begin{document}

\tableofcontents

\section{Files}
Each file should start with the header provided in \inlinecode{misc/HEADER}.
Specify the author(s) and add a description of the content of the file.
If you make significant changes to an existing file, add yourself to the list of authors in the header.

\subsection{File organization}
Each class should be defined in a separate file whose name is the class name with upper case letters replaced by an underscore (\inlinecode{_}) and the corresponding lower case letter.
(Exception: If the upper case letter is the first letter, no underscore is prepended).

Header files should be placed in \verb|include/dca/|, source files in \verb|src/|.
Tests should be placed according to their type in \verb|test/unit/|, \verb|test/integration/|, \verb|test/system-level/| and \verb|test/performance/|, respectively.
\verb|include/dca/|, \verb|src/| and \verb|test/unit/| should have the same tree structure.

\subsection{File Names}
Filenames should be all lowercase and can include numbers and underscores (\inlinecode{_}).
Uppercase letters are allowed for abbreviations and chemical compounds.
Header files use the extension \verb|.hpp|, while source files use the extension \verb|.cpp|.

The filenames of tests are composed by the filename of the object tested and the suffix \inlinecode{_test}.
The filenames of \emph{fake} and \emph{mock} objects used in tests are composed by the prefixes \inlinecode{fake_} and \inlinecode{mock_}, respectively, and the filename of the object that is imitated.

\subsection{Header files}
All header files should be self-contained.
This means that each header can be included without defining anything special.
In particular each header file should include all the headers needed and have \inlinecode{\#define} guards to prevent multiple inclusion.

If a template or inline function is declared in a \verb|.hpp| file, the function has to be defined in the same file.

The symbol name of the \inlinecode{\#define} guards should be \inlinecode{PATH_FILENAME_HPP}, where \inlinecode{PATH} is based on the \verb|include| directory.
For example, the file\\ \verb|DCA/include/dca/foo/bar.hpp| should have the following guard:

\begin{lstlisting}
#ifndef DCA_FOO_BAR_HPP
#define DCA_FOO_BAR_HPP
...
#endif  // DCA_FOO_BAR_HPP
\end{lstlisting}

\subsection{Includes}
Header files should be included with the full path based on the \verb|include| directory.
For example, the file \verb|DCA/include/dca/foo/bar.hpp| should be included as
\begin{lstlisting}
#include "dca/foo/bar.hpp"
\end{lstlisting}
Even if the included file is located in the same directory as the including file this rule should be obeyed. \\

For readability it is suggested to use the standard order of includes:
\begin{enumerate}
	\item Related header
	\item C library headers
	\item C++ library headers
	\item Other libraries' headers
	\item DCA++ headers
\end{enumerate}
In each section the included files should be sorted in alphabetical order.
\section{Naming}

The names should be as descriptive as possible.

\subsection{Type and Class Names}

Type and class names should start with a capital letter and have a capital letter for each new word.
Underscores (\inlinecode{_}) are not allowed.

The \inlinecode{using} syntax is preferred to \inlinecode{typedef} for type definitions.

\subsection{Namespace Names}

Namespace names should be one word, lowercase.

\subsection{Variable Names}

\subsubsection{Variable Names}

Variable names should contain only lowercase characters and underscores (\inlinecode{_}).
The underscore is not allowed as first or last character.

\subsubsection{Class Data Members}

Class data members names follows the convention of variable names with a trailing underscore (\inlinecode{_}).

\subsection{Physics Functions and Domains}

In order to be consistent with the usual mathematical notation, variables and types representing physics functions and domains are allowed to deviate from the conventions defined above.

TODO: Fix convention for domains.

\subsection{Function Names}

Function names should start with a lowercase character and have a capital letter for each new word.
Underscores (\inlinecode{_}) are not allowed.

The only exception to this rule are getter and setter methods for class data members.
They follow the naming convention \inlinecode{set_variable} and \inlinecode{get_variable}, respectively (see~\ref{subsec:class_format}).

\subsection{Lambda Expressions}

Named lambda expressions follow the naming convention for variables:

\begin{lstlisting}[showspaces=false]
auto my_lambda = [](int i) { return i + 4; };
\end{lstlisting}

\subsection{Macro Names}

Macro names should be all uppercase and can include underscores (\inlinecode{_}).
The underscore is not allowed as first or last character.

\subsection{Test Case and Test Names}

Test case and test names should start with a capital letter and have a capital letter for each new word.
\emph{Test case} names should end with \inlinecode{Test}.

\begin{lstlisting}[showspaces=false]
TEST(MatrixCPUTest, ElementAccess) {
    ...
}
\end{lstlisting}


\section{Comments}

\subsection{Comment Style}

Use the \inlinecode{// Comment} syntax.
The \inlinecode{/* ... */} syntax is allowed only for unused variables in function declarations.

\subsection{Function Comments}

Each function whose name does not fully describe the operations performed should have a comment block that describes what the function does.

For function parameters whose type is non-const reference or pointer to non-const memory,
it should be specified if they are input (In:), output (Out:) or input-output parameters (InOut:).

Example:
\begin{lstlisting}
// Updates foo and computes bar using in_1 .. in_5.
// In: in_3, in_5
// In/Out: foo
// Out: bar
void computeFooBar(Type in_1, const Type& in_2, Type& in_3,
                   const Type* in_4, Type* in_5, Type& foo,
                   Type& bar);
\end{lstlisting}

\subsection{Variable Comments}

No comments, since the name should be self-descriptive.

\subsection{Class Comments}

Every class should have a short description (in the header of the file) of what it is and what is does.
Comments for public class member functions follow the same rules as general function comments.
Comments for private members are allowed, but not mandatory.

\subsection{Implementation Comments}

Tricky, complicated or important code blocks should have comments before them.
Line comments should be separated from the code by 2 spaces.
If multiple subsequent lines have a comment they should be aligned.

Example:
\begin{lstlisting}
statement;        // Comment here so the comments line up.
longerstatement;  // Two spaces between code and comment.
\end{lstlisting}

If you modify a piece of code, also adapt the comments that belong to it if necessary.

\subsection{Comments excluded in releases}
[TODO: Review this section.]\\
Comments formatted as \inlinecode{// TODO: Comment}, \inlinecode{// INTERNAL: Comment} and \inlinecode{// REVIEW: Comment} are removed in the releases.

\subsection{Punctuation, Spelling and Grammar}

Pay attention to correct punctuation, spelling, and grammar. This will improve the clarity and readability of the comments.

\section{Formatting}

Use the provided clang-format style to format \verb|.hpp| and \verb|.cpp| files.

\subsection{Line Length}

The length of each line of your code should, in principle, be at most \emph{100} characters.
This limit can be exceeded by few characters in special cases.

\subsection{Scopes}

Do not use scopes for formatting reason.

\subsection{Horizontal Spacing}

No trailing whitespaces should be added to any line.
Use no space before a comma (\inlinecode{,}) and a semicolon (\inlinecode{;}) and add a space after them if they are not at the end of a line.

\subsubsection{Parenthesis}

Parenthesis should have no internal padding, and one space external padding.
The one space external padding is not allowed in the following cases:
\begin{itemize} \itemsep -3pt
 \vspace{-2ex}
 \item between two opening of two closing parenthesis,
 \item between the function name and its arguments,
 \item between a closing parenthesis and a comma (\inlinecode{,}) or a semicolon (\inlinecode{;}).
\end{itemize}

\subsubsection{Binary Operators}

The assignment operator should always have spaces around it.
Other operators may have spaces around them, but it is not mandatory.

\subsubsection{Unary Operators}

Do not put any space between an unary operator and their argument.

\subsubsection{Types}

The angle brackets of the templates should not have any external and internal padding.
(However, for \texttt{C++03} codes the space between two closing angle brackets is mandatory.)

Examples:
\begin{lstlisting}
Type* var;
Type& var;

Class1<Class2<type1>> object;  // OK with C++11, does not
                               // compile with C++03.
\end{lstlisting}

\subsection{Vertical Spacing}

Use empty lines when it helps to improve the readability of the code, but do not use too many.
Do not use empty lines after a brace which opens a scope,
or before a brace which closes a scope.
Each file should contain an empty line at the end of the file.
Some editors add an empty line automatically, some do not.

\subsection{Indentation}

Indentation consists of 2 spaces.
Do not use tabs in the code.

\subsection{Variable Declarations and Definitions}

\begin{itemize}

\item Avoid declaring multiple variables in the same declaration, especially if they are not fundamental types:

\begin{lstlisting}[showspaces=false]
int x, y;                        // Not recommended.
Matrix a("my-matrix"), b(size);  // Disallowed.

// Preferred way.
int x;
int y;
Matrix a("my-matrix");
Matrix b(10);
\end{lstlisting}

\item Use the following order for keywords and modifiers in  variable declarations:

\begin{lstlisting}[showspaces=false]
// General type
[static] [const/constexpr] Type variable_name;

// Pointer
[static] [const] Type* [const] variable_name;

// Integer
// int is optional, if a signedness or size modifier is present.
[static] [const/constexpr] [signedness] [size] int variable_name;

// Examples:
static const Matrix a(10);
const double* const d(3.14);
constexpr unsigned long l(42);
\end{lstlisting}

\end{itemize}

\subsection{Function Declarations and Definitions}

The return type should be on the same line as the function name.
Parameters should be on the same line, too, unless they do not fit on it.
Include the parameter names also in the declaration of a function, i.e. avoid
\begin{lstlisting}
Type function(Type1, Type2, Type3);
\end{lstlisting}
In function declarations comment the unused parameter names,\\
\inlinecode{Type /* unused_parameter_name */}

Examples:
\begin{lstlisting}
Type Class::function(Type1 par1, Type2 par2) {
  statement;
  ...
}

Type LongNameClass::longNameFunction(Type1 par1, Type2 par2
                                     Type3 par3) {
  statement;
  ...
}
\end{lstlisting}

% For formatting purpose, can be removed.
\vspace{3em}

In case of a long list of parameters prefer
\begin{lstlisting}
Type LongNameClass::longNameFunction(
    Type1 long_name_par1, Type2 long_name_par2,Type3 par3) {
  statement;
  ...
}
\end{lstlisting}
to
\begin{lstlisting}
Type LongNameClass::longNameFunction(Type1 long_name_par1,
                                     Type2 long_name_par2,
                                     Type3 par3) {
  statement;
  ...
}
\end{lstlisting}

\subsection{Function Calls}

Write the call on a single line if the length of the line does not exceed the maximum limit.
If it does, wrap the arguments at the parenthesis or start the arguments on a new line using 4 spaces indent.
Use the method which uses the smaller amount of lines.

Examples:
\begin{lstlisting}
function(par1, par2, par3);

function(par1, par2,
         par3);

function(
    par1, par2, par3);
\end{lstlisting}

\subsection{Conditionals}

Examples:
\begin{lstlisting}
if (condition)
  statement;
else
  statement;

if (condition) {
  statement;
}
else if (condition2) {
  statement;
}
else {
  statement;
}
\end{lstlisting}

\subsection{Switch statement}

Switch statements should always have a default case.

Example:
\begin{lstlisting}
switch (var) {
  case 0:
    statement1;
    statement2;
    break;

  case 1:
    statement1;
    statement2;
    break;

  default:
    statement1;
    statement2;
}
\end{lstlisting}

\subsection{Loops}

Examples:
\begin{lstlisting}
for (statement; condition; statement)
  statement;

for (statement; condition; statement) {
  statement1;
  statement2;
}

while (condition)
  statement;

while (condition) {
  statement1;
  statement2;
}

do {
  statement;
}
while (condition);
\end{lstlisting}

\subsection{Preprocessor Directives}

The preprocessor directives are not indented.
The hash is the first character of the line.

\subsection{Class Format}
\label{subsec:class_format}

\inlinecode{public}, \inlinecode{protected} and \inlinecode{private} keywords are not indented.

Example:
\begin{lstlisting}
class Foo : public Bar {
public:
  Foo();
  explicit Foo(int var);

  void function();
  void emptyFunction() {}

  void set_var(const int var) {
    var_ = var;
  }
  int get_var() const {
    return var_;
  }

private:
  bool privateFunction();

  int var_;
  int var2_;
};
\end{lstlisting}

\subsubsection{Constructor Initializer Lists}

Examples:
\begin{lstlisting}
// When everything fits on one line:
Foo::Foo(int var) : var_(var) {
  statement;
}

// If the signature and the initializer list do not
// fit on one line, the colon is indented by 4 spaces:
Foo::Foo(int var)
    : var_(var), var2_(var + 1) {
  statement;
}

// If the initializer list occupies more lines,
// they are aligned in the following way:
Foo::Foo(int var)
    : some_var_(var),
      some_other_var_(var + 1) {
  statement;
}

// No statements:
Foo::Foo(int var)
    : some_var_(var) {}
\end{lstlisting}

\subsection{Namespace Formatting}

The content of a namespace is not indented. \\
A comment should indicate when a namespace is closed.
ClangFormat >= 6.0 does this automatically. \\
If nested namespaces are used, a comment with the full namespace is required after opening a set of namespaces or an inner namespace.

Examples:
\begin{lstlisting}
namespace ns{
void foo();
}  // namespace ns
\end{lstlisting}

\begin{lstlisting}
namespace ns1{
namespace ns2{
// ns1::ns2::
void foo();

namespace ns3{
// ns1::ns2::ns3::
void bar();
}  // namespace ns3
}  // namespace ns2

namespace ns4{
namespace ns5{
// ns1::ns4::ns5::
void foo();
}  // namespace ns5
}  // namespace ns4
}  // namespace ns1
\end{lstlisting}


\section{Other C++ Features}

\subsection{Pre-increment and pre-decrement}

Use the pre-increment (pre-decrement) operator when a variable is incremented (decremented) and the value of the expression is not used.
In particular, use the pre-increment (pre-decrement) operator for loop counters:

\begin{lstlisting}[showspaces=false]
for (int i = 0; i < N; ++i) {
  doSomething();
}
\end{lstlisting}

The post-increment and post-decrement operators create an unnecessary copy, that the compiler cannot optimize away in the case of iterators or other classes with overloaded increment and decrement operators.

\subsection{Alternative Operator Representations}

Alternative representations of operators and other tokens such as \inlinecode{and}, \inlinecode{or}, and \inlinecode{not} instead of \inlinecode{&&}, \inlinecode{||}, and \inlinecode{!} are not allowed.
For the reason of consistency, the far more common primary tokens should always be used.

\subsection{Use of const}

\begin{itemize}

\item Add the \inlinecode{const} qualifier to all function parameters that are not modified in the function body.
For parameters passed by value, only add the keyword in the function definition.

\begin{lstlisting}[showspaces=false]
// Declaration
int computeFoo(int bar, const Matrix& m) {

// Definition
int computeFoo(const int bar, const Matrix& m) {
  int foo = 42;

  // Compute foo without changing bar or m.
  // ...
	
  return foo;
}
\end{lstlisting}

\end{itemize}

\subsection{In-class member initialization}

In-class member initialization is recommended for members that are \emph{always} initialized with the same value to avoid code duplication.
In all other cases it is disallowed.

\begin{lstlisting}[showspaces=false]
std::string getNameFromId(const int id);

class Customer {
public:
  Customer(const std::string& name = "Doe",
           const std::string& address = default_address_)
    : name_(name), address_(address) {}
  Customer(const int id,
           const std::string& address = default_address_)
    : name_(getNameFromId(id)), address_(address) {}

private:
  static const std::string default_address_;

  std::string name_;
  std::string address_;
  int num_orders_ = 0;
};

const std::string Customer::default_address_ = "Zurich";
\end{lstlisting}

\end{document}
